### 2023-01-03 editorial-communication/response-to-the-reviewers.Qmd

---
title: "Response to the reviewers"
date: today
date-format: long
format:
  html:
    theme: cosmo 
    highlight-style: tango
    embed-resources: true
    number-sections: true
---

### 2023-01-03 editorial-communication/response-to-the-reviewers.Qmd

---
title: "Response to the reviewers"
date: today
date-format: long
format:
  docx: 
    reference-doc: response-template.docx
  html:
    theme: cosmo 
    highlight-style: tango
    embed-resources: true
    number-sections: true
---

### 2023-01-30 manuscript.tex
Merge pull request #303 from SORTEE-Github-Hackathon/r1-merge [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/b2e5480009b1ff9b626369295a0af0b08b0cf535.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/b2e5480009b1ff9b626369295a0af0b08b0cf535/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/4045188171

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/077be9d884d5bc703a0a7db131e90f2d35254fbc/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/077be9d884d5bc703a0a7db131e90f2d35254fbc}{SORTEE-Github-Hackathon/manuscript@077be9d}
on January 2, 2023.

@@

Despite these benefits, the use of GitHub in ecology and evolution is not widespread.
To help researchers in ecology and evolution adopt useful features from GitHub in their own workflows, we review thirteen practical ways to use the platform.
We outline features ranging from low to high technical difficulty: storing code, managing projects, coding collaboratively, conducting peer review, and writing a manuscript.
Given that members of a research team may have different technical skills and responsibilities, we describe how the optimal use of GitHub features may vary among members of a research collaboration.
As more ecologists and evolutionary biologists establish their workflows using GitHub, the field can continue to push the boundaries of collaborative, transparent, and open research.

@@

This paper is the result of an academic hackathon held during the 2021 conference for the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE, \url{https://www.sortee.org}).
We convened around thirty researchers in EEB with varying levels of familiarity with the usage of GitHub in research projects to showcase and discuss how existing features can contribute to documentation and collaboration in EEB research.
During the hackathon, we identified the need for formal discussions on how EEB researchers can benefit from GitHub and its features.
Here, we outline thirteen practical ways that EEB researchers can use GitHub features for more collaborative, transparent, and reproducible science.
We also provide critical perspectives on features that could be improved and catered towards research development.

@@

\hypertarget{thirteen-practical-ways-github-can-accelerate-research-in-ecology-and-evolution}{%
\subsection{Thirteen practical ways GitHub can accelerate research in ecology and evolution}\label{thirteen-practical-ways-github-can-accelerate-research-in-ecology-and-evolution}}

@@

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having copies of these files safely stored is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub to store (or backup) their research compendium (\protect\hyperlink{ref-MwwMapRG}{Marwick et al., 2018}) to a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving services, allowing file versions to be accessed and restored, and contributes to open science (see sections below).
While some EEB researchers may prefer to store their code in Google Drive or DropBox, the process of sharing code through these services becomes cumbersome when collaborators have to download and re-upload edited code.
On GitHub, code is by default public, and the process of forking and submitting pull requests (see \protect\hyperlink{definitions}{Box 1}) ensures a code owner has full control over code and documents in a repository.
The ability to make a GitHub repository private is particularly advantageous for storing materials without necessarily sacrificing privacy, which may be preferred or even essential for certain compendia.
Researchers can then grant specific collaborators read and/or write access to private repositories to pursue pre-publication analyses or writing in privacy, or even to collaborate on research that will permanently remain private.

@@

\hypertarget{virtual-laboratory-notebooks}{%
\subsubsection{Virtual laboratory notebooks}\label{virtual-laboratory-notebooks}}

Laboratory notebooks help researchers track their research notes, methods, policies and protocols (\protect\hyperlink{ref-wwHxTOtm}{Kanza et al., 2017}).
Virtual laboratory notebooks can be stored on GitHub and provide the benefits of simultaneous, centralized and selective access, and allows for the easy update of policies and experiment protocols (\protect\hyperlink{ref-10V7x4H4l}{Schnell, 2015}).
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory notebooks across many fields (\protect\hyperlink{ref-10ghgV3S8}{J. Perkel, 2016}), including applied ecology (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}), biogeography and global change biology (\emph{e.g.}, \url{https://github.com/HuckleyLab/how_we_work}), and microbial ecology (\emph{e.g.}, \url{https://github.com/CarBBAS/uqam-guide}).

At least for aspects of a research project that involve writing code, a GitHub repository is a form of a laboratory notebook; when changes are made to files in a version-controlled repository, the author of those changes makes a commit (\protect\hyperlink{definitions}{Box 1}) accompanied by a description of changes.
Later, the entire history of commits and their commit messages are viewable and can be audited similar to a physical laboratory notebook (\protect\hyperlink{ref-4ny1onB0}{Ram, 2013}).
GitHub issues (see Project management section below) can be used to prioritize laboratory objectives and goals, as well as track any progress updates.

@@

In academic settings, GitHub can serve as a platform for everyone (advisors, advisees, collaborators) to share in-progress work, and flag specific challenges or questions for each other (\protect\hyperlink{tbl:roles}{Table 2}).
Periodic code review (\protect\hyperlink{ref-hm9PaCLD}{Song et al., 2020}) is useful for identifying errors early in the scientific process, and informing further training and mentorship to fill gaps in skills.
This is facilitated by a group of core features of git and GitHub that allow contributors to work simultaneously on the same project without interfering with each other's work.
``Branching'' and ``forking'' (see \protect\hyperlink{definitions}{Box 1}) allow users to work independently on a linked copy of a repository and then later merge their changes into the main project.
When a set of related changes contributing some new feature or milestone are ready to be evaluated and merged, one makes a ``pull request'' (see \protect\hyperlink{definitions}{Box 1}).
Pull requests are a formalized way of integrating changes to avoid conflicts that arise when users make simultaneous edits to the same parts of files (\protect\hyperlink{ref-1BJcvyTmV}{Vale et al., 2019}).
The GitHub interface for a pull request allows collaborators to comment on specific lines of code/text being changed and even to provide suggestions that the pull request author can implement with the click of a button, greatly facilitating peer review.
Moreover, version control (\protect\hyperlink{ref-K7nbP1Ty}{Crystal-Ornelas et al., 2022}) allows researchers to make changes in code or documents without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.

@@

Projects in ecology and evolution often involve graduate students, research assistants, and post-doctoral fellows, who hold limited-term positions (\protect\hyperlink{ref-D4C4k4ak}{Fehr et al., 2021}).
Without clear plans on project continuity, the research code and data management upkeep tends to fall off as researchers move on to new projects or to other institutions.
Additionally, code and data can be difficult to access when kept only on personal devices (\protect\hyperlink{ref-19kmNxiHc}{Vines et al., 2014}).

GitHub can facilitate project continuity among research code by making code handover easier (\protect\hyperlink{ref-D4C4k4ak}{Fehr et al., 2021}; \protect\hyperlink{ref-4ny1onB0}{Ram, 2013}).
Through version control, the history of code and data from projects in ecology and evolution can be tracked accessible to future laboratory members and collaborators (\protect\hyperlink{ref-3DKwn1sY}{Lowndes et al., 2017}).
Repositories and organizations can have designated data and code owners {[}or more appropriate, stewards; see Hampton et al. (\protect\hyperlink{ref-iIEKCTLU}{2015}), \emph{About Code Owners} (\protect\hyperlink{ref-s91uGRZ2}{n.d.}){]}, who can also change through time allowing for the transition of code between research cohorts.
Other project collaborators can contribute to repository design and development, and their active involvement can both aid authors ability to act as guarantors of the project, and the clarity and reproducibility of the project for future users.
In (Figure \ref{fig:github-diagram}), we highlight several elements of good repository structure, and the various ways that contributors may interact with them.

@@

\hypertarget{organizing-and-managing-teams}{%
\subsubsection{Organizing and managing teams}\label{organizing-and-managing-teams}}

@@

The thirteen use cases we described here can leverage GitHub to enable more transparent and collaborative research in ecology and evolution (Figure \ref{fig:scatterblob}).
Most of these uses can be quickly integrated into the research practices of users (\emph{e.g.}, storing data, creating virtual notebooks, making code citable).
Making stored data and code citable usually involves creating a repository on GitHub, pushing code and data, and then integrating a DOI minting service to the repository (\emph{e.g.}, with Zenodo; see below).
On the other hand, some examples we described here, including course material development, web hosting, and automation, require greater effort and time commitment, but have the potential to make EEB research more open, accessible, and collaborative.
Managing full research projects or laboratories on GitHub require careful thought as to how to delegate tasks such as reviewing pull requests or creating issues.
For example, collaboratively authoring a paper using GitHub, as we have done here, involves a learning curve for co-authors less familiar with the intricacies of GitHub, requires overhead to set up the automation framework through GitHub Actions, and especially, the commitment from all co-authors to use GitHub when modifying and reviewing the text.
Despite the potential applications of GitHub to EEB research, we acknowledge that researchers might still look to other platforms for research collaboration.

@@

We describe thirteen practical ways that ecologists and evolutionary biologists can adopt GitHub to improve their research workflow and make it more open and reproducible.
We provide definitions (\protect\hyperlink{definitions}{Box 1}) and types of users (Figure \ref{fig:github-diagram}) to help researchers identify and prioritize the skills and tools to learn and apply.
We highlight tools providing high collaborative potential (\emph{e.g.}, open science discussion, collaborative code editing) to more individual focused (\emph{e.g.}, storing code and data, building a website).
We argue that the tools readily available in GitHub have the potential to make ecology and evolution more open, reproducible and transparent.
With this comprehensive review of how EEB researchers can use GitHub, we encourage researchers at any career stage to adopt GitHub as a platform for sharing and collaboration.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-s91uGRZ2}{}}%
\emph{About code owners}. (n.d.). GitHub Docs. Retrieved January 2, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/repositories/managing-your-repositorys-settings-and-features/customizing-your-repository/about-code-owners}

\leavevmode\vadjust pre{\hypertarget{ref-1Co6ZZjF1}{}}%
\emph{About large files on GitHub}. (n.d.). GitHub Docs. Retrieved January 2, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/repositories/working-with-files/managing-large-files/about-large-files-on-github}

\leavevmode\vadjust pre{\hypertarget{ref-RhBKe0MG}{}}%
\emph{About Projects}. (n.d.). GitHub Docs. Retrieved January 2, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/issues/planning-and-tracking-with-projects/learning-about-projects/about-projects}

\leavevmode\vadjust pre{\hypertarget{ref-TOsASkn5}{}}%
\emph{Adding a license to a repository}. (n.d.). GitHub Docs. Retrieved January 2, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/communities/setting-up-your-project-for-healthy-contributions/adding-a-license-to-a-repository}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-6CMMeSeD}{}}%
Bryan, J., \& TAs, T. S. 545. (n.d.). \emph{STAT 545}. Retrieved January 2, 2023, from \url{https://stat545.com/}

@@

\leavevmode\vadjust pre{\hypertarget{ref-lx49NGto}{}}%
\emph{Connect GitHub to a Project - OSF Support}. (n.d.). Retrieved January 2, 2023, from \url{https://help.osf.io/article/211-connect-github-to-a-project}

\leavevmode\vadjust pre{\hypertarget{ref-T03Api6e}{}}%
\emph{Continuous Integration and Delivery}. (n.d.). CircleCI. Retrieved January 2, 2023, from \url{https://circleci.com/}

\leavevmode\vadjust pre{\hypertarget{ref-K7nbP1Ty}{}}%
Crystal-Ornelas, R., Edwards, B. P. M., Hébert, K., Hudgins, E. J., Reyes, L. L. S., Scott, E. R., Grainger, M. J., Foroughirad, V., Binley, A. D., Brookson, C. B., Gaynor, K. M., Sabet, S. S., Güncan, A., Hillemann, F., Weierbach, H., Gomes, D. G. E., \& Braga, P. H. P. (2022). \emph{Not just for programmers: How GitHub can accelerate collaborative and reproducible research in ecology and evolution}. Manubot. \url{https://SORTEE-Github-Hackathon.github.io/manuscript/}

@@

\leavevmode\vadjust pre{\hypertarget{ref-NUXbp429}{}}%
\emph{Features • GitHub Actions}. (n.d.). GitHub. Retrieved January 2, 2023, from \url{https://github.com/features/actions}

@@

\leavevmode\vadjust pre{\hypertarget{ref-1CM2EcdVk}{}}%
\emph{Git - A Short History of Git}. (n.d.). Retrieved January 2, 2023, from \url{https://git-scm.com/book/en/v2/Getting-Started-A-Short-History-of-Git}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ZvrOcg9w}{}}%
Hester, J. B., the STAT 545 TAs, Jim. (n.d.). \emph{Let's Git started \textbar{} Happy Git and GitHub for the useR}. Retrieved January 2, 2023, from \url{https://happygitwithr.com/}

@@

\leavevmode\vadjust pre{\hypertarget{ref-3UAritXO}{}}%
\emph{Issues · fmsabatini/sPlotOpen\_Manuscript}. (n.d.). GitHub. Retrieved January 2, 2023, from \url{https://github.com/fmsabatini/sPlotOpen_Manuscript}

@@

\leavevmode\vadjust pre{\hypertarget{ref-wwHxTOtm}{}}%
Kanza, S., Willoughby, C., Gibbins, N., Whitby, R., Frey, J. G., Erjavec, J., Zupančič, K., Hren, M., \& Kovač, K. (2017). Electronic lab notebooks: can they replace paper? \emph{Journal of Cheminformatics}, \emph{9}(1). \url{https://doi.org/10.1186/s13321-017-0221-3}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Re6Eg2va}{}}%
\emph{Manubot - Manuscripts, open and automated}. (n.d.). Retrieved January 2, 2023, from \url{https://manubot.org}

@@

\leavevmode\vadjust pre{\hypertarget{ref-10V7x4H4l}{}}%
Schnell, S. (2015). Ten Simple Rules for a Computational Biologist's Laboratory Notebook. \emph{PLOS Computational Biology}, \emph{11}(9), e1004385. \url{https://doi.org/10.1371/journal.pcbi.1004385}

@@

\leavevmode\vadjust pre{\hypertarget{ref-hm9PaCLD}{}}%
Song, X., Goldstein, S. C., \& Sakr, M. (2020, June 2). Using Peer Code Review as an Educational Tool. \emph{Proceedings of the 2020 ACM Conference on Innovation and Technology in Computer Science Education}. ITiCSE '20: Innovation and Technology in Computer Science Education. \url{https://doi.org/10.1145/3341525.3387370}

@@

\leavevmode\vadjust pre{\hypertarget{ref-oGbKxHzj}{}}%
\emph{Using Git and GitHub}. (n.d.). Retrieved January 2, 2023, from \url{https://www.overleaf.com/learn/how-to/Using_Git_and_GitHub}

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/b2e5480009b1ff9b626369295a0af0b08b0cf535/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/b2e5480009b1ff9b626369295a0af0b08b0cf535}{SORTEE-Github-Hackathon/manuscript@b2e5480}
on January 30, 2023.

@@

Despite these benefits, the use of GitHub in ecology and evolution is not widespread.
To help researchers in ecology and evolution adopt useful features from GitHub in their own workflows, we review twelve practical ways to use the platform.
We outline features ranging from low to high technical difficulty: storing code, managing projects, coding collaboratively, conducting peer review, and writing a manuscript.
Given that members of a research team may have different technical skills and responsibilities, we describe how the optimal use of GitHub features may vary among members of a research collaboration.
As more ecologists and evolutionary biologists establish their workflows using GitHub, the field can continue to push the boundaries of collaborative, transparent, and open research.

@@

This paper is the result of an academic hackathon held during the 2021 conference for the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE, \url{https://www.sortee.org}).
We convened around thirty researchers in EEB with varying levels of familiarity with the usage of GitHub in research projects to showcase and discuss how existing features can contribute to documentation and collaboration in EEB research.
During the hackathon, we identified the need for formal discussions on how EEB researchers can benefit from GitHub and its features.
Here, we outline twelve practical ways that EEB researchers can use GitHub features for more collaborative, transparent, and reproducible science.
We also provide critical perspectives on features that could be improved and catered towards research development.

@@

\hypertarget{twelve-practical-ways-github-can-accelerate-research-in-ecology-and-evolution}{%
\subsection{Twelve practical ways GitHub can accelerate research in ecology and evolution}\label{twelve-practical-ways-github-can-accelerate-research-in-ecology-and-evolution}}

@@

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Safely storing these files is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub to backup their research compendium (\protect\hyperlink{ref-MwwMapRG}{Marwick et al., 2018}) into a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving, allowing file versions to be accessed and restored, and further contributing to open science practices (Borghi \& Van Gulick (\protect\hyperlink{ref-gLby7jt1}{2022})).
Every change made to files in a version-controlled repository accompany a description of modifications and are authored (\protect\hyperlink{definitions}{Box 1}).
Later, the entire history of commits and their commit messages are viewable and can be audited similar to a physical laboratory notebook (\protect\hyperlink{ref-4ny1onB0}{Ram, 2013}).
On GitHub, code can be made public, and the process of forking and submitting pull requests (see \protect\hyperlink{definitions}{Box 1}) ensures code owners have full control over code and documents in a repository.
The ability to make a GitHub repository private is particularly advantageous for storing materials without necessarily sacrificing privacy, which may be preferred or even essential for certain compendia.
Researchers can then grant specific collaborators read and/or write access to private repositories to pursue pre-publication analyses or writing in privacy, or even to collaborate on research that will permanently remain private.

@@

Projects in ecology and evolution often involve research professionals holding limited-term positions, such as graduate students, research assistants, and post-doctoral fellows (\protect\hyperlink{ref-D4C4k4ak}{Fehr et al., 2021}).
Without clear plans on project continuity, the research code and data management upkeep tend to fall off as researchers move on to new projects or other institutions.
Additionally, code and data can be difficult to access when kept only on personal devices (\protect\hyperlink{ref-19kmNxiHc}{Vines et al., 2014}).

GitHub can facilitate project continuity among research code by making code handover easier (\protect\hyperlink{ref-D4C4k4ak}{Fehr et al., 2021}; \protect\hyperlink{ref-4ny1onB0}{Ram, 2013}).
Through version control, the history of code and data from projects in ecology and evolution can be tracked accessible to future laboratory members and collaborators (\protect\hyperlink{ref-3DKwn1sY}{Lowndes et al., 2017}).
Repositories and organizations can have designated data and code owners {[}or more appropriate, stewards; see Hampton et al. (\protect\hyperlink{ref-iIEKCTLU}{2015}), \emph{About Code Owners} (\protect\hyperlink{ref-s91uGRZ2}{n.d.}){]}, who can also change through time allowing for the transition of code between research cohorts (see also \protect\hyperlink{organizations}{``Organizing and managing teams''}).
Other project collaborators can contribute to repository design and development, and their active involvement can both aid authors ability to act as guarantors of the project, and the clarity and reproducibility of the project for future users.
In (Figure \ref{fig:github-diagram}), we highlight several elements of good repository structure, and the various ways that contributors may interact with them.

GitHub can facilitate lab continuity by being used as a platform for storing, selectively sharing, and easily updating lab policies and experimental protocols.
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory handbooks across many fields (\protect\hyperlink{ref-10ghgV3S8}{J. Perkel, 2016}), including applied ecology (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}), biogeography and global change biology (\emph{e.g.}, \url{https://github.com/HuckleyLab/how_we_work}), and microbial ecology (\emph{e.g.}, \url{https://github.com/CarBBAS/uqam-guide}).

@@

In academic settings, GitHub can serve as a platform for everyone (advisors, advisees, collaborators) to share in-progress work, and flag specific challenges or questions for each other (\protect\hyperlink{tbl:roles}{Table 2}).
Periodic code review (\protect\hyperlink{ref-hm9PaCLD}{Song et al., 2020}) is useful for identifying errors early in the scientific process, and informing further training and mentorship to fill gaps in skills.
This is facilitated by a group of core features of git and GitHub that allow contributors to work simultaneously on the same project without interfering with each other's work.
``Branching'' and ``forking'' (see \protect\hyperlink{definitions}{Box 1}) allow users to work independently on a linked copy of a repository and then later merge their changes into the main project.
When a set of related changes contributing some new feature or milestone are ready to be evaluated and merged, one makes a ``pull request'' (see \protect\hyperlink{definitions}{Box 1}).
Pull requests are a formalized way of integrating changes to avoid conflicts that arise when users make simultaneous edits to the same parts of files (\protect\hyperlink{ref-1BJcvyTmV}{Vale et al., 2019}).
The GitHub interface for a pull request allows collaborators to comment on specific lines of code/text being changed and even to provide suggestions that the pull request author can implement with the click of a button, greatly facilitating peer review.
Moreover, version control (\protect\hyperlink{ref-K7nbP1Ty}{Crystal-Ornelas et al., 2023}) allows researchers to make changes in code or documents without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.

@@

@@

\hypertarget{organizations}{%
\subsubsection{Organizing and managing teams}\label{organizations}}

@@

The twelve use cases we described here can leverage GitHub to enable more transparent and collaborative research in ecology and evolution (Figure \ref{fig:scatterblob}).
Most of these uses can be quickly integrated into the research practices of users (\emph{e.g.}, storing data, creating virtual notebooks, making code citable).
Making stored data and code citable usually involves creating a repository on GitHub, pushing code and data, and then integrating a DOI minting service to the repository (\emph{e.g.}, with Zenodo; see below).
On the other hand, some examples we described here, including course material development, web hosting, and automation, require greater effort and time commitment, but have the potential to make EEB research more open, accessible, and collaborative.
Managing full research projects or laboratories on GitHub require careful thought as to how to delegate tasks such as reviewing pull requests or creating issues.
For example, collaboratively authoring a paper using GitHub, as we have done here, involves a learning curve for co-authors less familiar with the intricacies of GitHub, requires overhead to set up the automation framework through GitHub Actions, and especially, the commitment from all co-authors to use GitHub when modifying and reviewing the text.
Despite the potential applications of GitHub to EEB research, we acknowledge that researchers might still look to other platforms for research collaboration.

@@

We describe twelve practical ways that ecologists and evolutionary biologists can adopt GitHub to improve their research workflow and make it more open and reproducible.
We provide definitions (\protect\hyperlink{definitions}{Box 1}) and types of users (Figure \ref{fig:github-diagram}) to help researchers identify and prioritize the skills and tools to learn and apply.
We highlight tools providing high collaborative potential (\emph{e.g.}, open science discussion, collaborative code editing) to more individual focused (\emph{e.g.}, storing code and data, building a website).
We argue that the tools readily available in GitHub have the potential to make ecology and evolution more open, reproducible and transparent.
With this comprehensive review of how EEB researchers can use GitHub, we encourage researchers at any career stage to adopt GitHub as a platform for sharing and collaboration.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-s91uGRZ2}{}}%
\emph{About code owners}. (n.d.). GitHub Docs. Retrieved January 30, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/repositories/managing-your-repositorys-settings-and-features/customizing-your-repository/about-code-owners}

\leavevmode\vadjust pre{\hypertarget{ref-1Co6ZZjF1}{}}%
\emph{About large files on GitHub}. (n.d.). GitHub Docs. Retrieved January 30, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/repositories/working-with-files/managing-large-files/about-large-files-on-github}

\leavevmode\vadjust pre{\hypertarget{ref-RhBKe0MG}{}}%
\emph{About Projects}. (n.d.). GitHub Docs. Retrieved January 30, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/issues/planning-and-tracking-with-projects/learning-about-projects/about-projects}

\leavevmode\vadjust pre{\hypertarget{ref-TOsASkn5}{}}%
\emph{Adding a license to a repository}. (n.d.). GitHub Docs. Retrieved January 30, 2023, from \url{https://ghdocs-prod.azurewebsites.net/en/communities/setting-up-your-project-for-healthy-contributions/adding-a-license-to-a-repository}

@@

\leavevmode\vadjust pre{\hypertarget{ref-gLby7jt1}{}}%
Borghi, J., \& Van Gulick, A. (2022). Promoting Open Science Through Research Data Management. \emph{Harvard Data Science Review}. \url{https://doi.org/10.1162/99608f92.9497f68e}

@@

\leavevmode\vadjust pre{\hypertarget{ref-6CMMeSeD}{}}%
Bryan, J., \& TAs, T. S. 545. (n.d.). \emph{STAT 545}. Retrieved January 30, 2023, from \url{https://stat545.com/}

@@

\leavevmode\vadjust pre{\hypertarget{ref-lx49NGto}{}}%
\emph{Connect GitHub to a Project - OSF Support}. (n.d.). Retrieved January 30, 2023, from \url{https://help.osf.io/article/211-connect-github-to-a-project}

\leavevmode\vadjust pre{\hypertarget{ref-T03Api6e}{}}%
\emph{Continuous Integration and Delivery}. (n.d.). CircleCI. Retrieved January 30, 2023, from \url{https://circleci.com/}

\leavevmode\vadjust pre{\hypertarget{ref-K7nbP1Ty}{}}%
Crystal-Ornelas, R., Edwards, B. P. M., Hébert, K., Hudgins, E. J., Reyes, L. L. S., Scott, E. R., Grainger, M. J., Foroughirad, V., Binley, A. D., Brookson, C. B., Gaynor, K. M., Sabet, S. S., Güncan, A., Hillemann, F., Weierbach, H., Gomes, D. G. E., \& Braga, P. H. P. (2023). \emph{Not just for programmers: How GitHub can accelerate collaborative and reproducible research in ecology and evolution}. Manubot. \url{https://SORTEE-Github-Hackathon.github.io/manuscript/}

@@

\leavevmode\vadjust pre{\hypertarget{ref-NUXbp429}{}}%
\emph{Features • GitHub Actions}. (n.d.). GitHub. Retrieved January 30, 2023, from \url{https://github.com/features/actions}

@@

\leavevmode\vadjust pre{\hypertarget{ref-1CM2EcdVk}{}}%
\emph{Git - A Short History of Git}. (n.d.). Retrieved January 30, 2023, from \url{https://git-scm.com/book/en/v2/Getting-Started-A-Short-History-of-Git}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ZvrOcg9w}{}}%
Hester, J. B., the STAT 545 TAs, Jim. (n.d.). \emph{Let's Git started \textbar{} Happy Git and GitHub for the useR}. Retrieved January 30, 2023, from \url{https://happygitwithr.com/}

@@

\leavevmode\vadjust pre{\hypertarget{ref-3UAritXO}{}}%
\emph{Issues · fmsabatini/sPlotOpen\_Manuscript}. (n.d.). GitHub. Retrieved January 30, 2023, from \url{https://github.com/fmsabatini/sPlotOpen_Manuscript}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Re6Eg2va}{}}%
\emph{Manubot - Manuscripts, open and automated}. (n.d.). Retrieved January 30, 2023, from \url{https://manubot.org}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-hm9PaCLD}{}}%
Song, X., Goldstein, S. C., \& Sakr, M. (2020, June 15). Using Peer Code Review as an Educational Tool. \emph{Proceedings of the 2020 ACM Conference on Innovation and Technology in Computer Science Education}. ITiCSE '20: Innovation and Technology in Computer Science Education. \url{https://doi.org/10.1145/3341525.3387370}

@@

\leavevmode\vadjust pre{\hypertarget{ref-oGbKxHzj}{}}%
\emph{Using Git and GitHub}. (n.d.). Retrieved January 30, 2023, from \url{https://www.overleaf.com/learn/how-to/Using_Git_and_GitHub}
